Damage localization based on vibrations

### 2017-03-31 rapport/paper.tex
MAJ paper

The purpose of this work is the analysis of a particular data-driven damage localization method,
 namely the FRF interpolation method \cite{dilena2015damage}. Our contribution is the development of a simulation to obtain acceleration measurement (which can run on several structures), a proposition of a variant of the FRF interpolation method, and the statistical analysis of the two methods.
This paper is organized as follows. In section II we  present the existing methods (model-based and data driven methods), 
then in section III we talk about the localization using FRF interpolation method, in section IV we detail the different models of vibrating structures used to obtain data and, finally, the experiences and its results in section V.

@@

@@

Figure (\ref{theo}) shows this loss of spatial regularity on the theoretical FRF, for a damage at the position 7.

@@

In this section, we suppose that we have $n$ sensors, which are located along an axis $z$, and we denote them by
$\{z_l\}, l\in \{1 ... n\}$. We also suppose we have $N$ measurements for each sensor.

For each sensor, we calculate the Frequency Response Function (FRF) from the acceleration's measures, at certain frequencies 
(which are $\frac{k}{N}, k \in \{0 ... N-1\}$).

@@

@@

The FRF correspond to the transfer function of the system. In frequency domain, if $F$ is the force that is apply on the system and $S$ is the response of this force, the FRF is given by:
\begin{equation}
H = \frac{S}{F}
\end{equation}

We want to calculate the FRF of the system, for each sensor localization. Here, we calculate an estimation of the FRF, called $H_1$ estimator \cite{cauberghe2004applied}. We suppose we have a vector of the forces $\{ f_k\}_{k=1}^N$ and a vector of the accelerations $\{ y_k\}_{k=1}^N$ for each sensor (the two vectors are size $N$). We need to be in frequency domain, so we calculate the discrete Fourier transform:
\begin{equation}
\begin{array}{ll}
F_n = \sum\limits_{k=1}^N f_k e^{-2i\pi kn/N} \ \ \ \ \ n \in \{1 ... N\} \\
Y_n = \sum\limits_{k=1}^N y_k e^{-2i\pi kn/N} \ \ \ \ \ n \in \{1 ... N\}
\end{array}
\end{equation}

@@

We have to calculate for each sensor $z_l$ and for each considered frequency $f_i$ the cubic polynomial spline interpolation of the transfer function $H_R(z_k,f_i)$.
The cubic spline interpolation is a piecewise continuous curve, passing through each of the values.
There is a separate cubic polynomial interpolation for each interval, each with its own coefficients.
Cubic polynomial spline interpolations produce an interpolated function that is continuous through to the second derivative.

@@

With this interpolation, we can calculate the error as it is described in the section IV.

\subsection{False and missing alarm}

@@

Thus, once have to determine a threshold $E_T$, such as $ \Delta E(z_l)$ must be taller than $E_T$ to have a high probability of identification of a damage at location $z_l$. Indeed, as we can see in the Figure (\ref{proba}), the curves which represent the fact to have a damage or not have a share area: if the interpolation error $ \Delta E(z_l)$ is in this area, one can commit two errors:
\begin{itemize}
\item if it is in the area $P_f(z_l)$ under the graph of $p_{E,0}(z_l)$, once can detect a damage since there isn't one (false alarm).
\item if it is in the area $P_m(z_l)$ under the graph of $p_{E,d}(z_l)$, once can not detect a damage since there is one (missing alarm).
\end{itemize}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/gaussiennes.png}
  \caption{Probabilities of false and missing alarm, from \cite{dilena2015damage}}
  \label{proba}
\end{figure}

@@

We propose here a variant of the FRF interpolation method. While we have worked on the implementation of the original method, we have observed the eigenvectors of the matrix $M^{-1}K$ (see part III), which have similar properties than the FRF. We noticed that the following method is better than the interpolation method (compute the interpolation error, then compute a difference) to detect damaged on the eigenvectors:
\begin{itemize}
\item first, compute the difference between the eigenvectors
\item second, compute the interpolation error on this difference
\end{itemize}

Figure(\ref{theodiff}) shows the difference betweem the two FRF, and we can see the loss of spatial regularity (damage at position 7).

@@

\subsection{Models of vibrating structures}% or Physical models

\subsubsection{Finite element model}

The dynamic behaviour of a mechanical system consisting of $n$ degrees of freedom (see Figure (\ref{finite_el})) is described by following matrix differential equation:
\begin{equation}
M \ddot{q}(t) + C \dot{q}(t) + K q(t) = f(t)
\label{fondamental}
\end{equation}
where $q(t)$ is the displacement vector at a continous time $t$, and $M, C, K \in \mathbb{R}^{n \times n}$, are the mass, damping and stiffness matrices.

\subsubsection{Continous-time state-space model} %(model parameters and reduction):

\subsubsection{Discrete-time state-space model} %(impulses response)

\subsection{Obtaining acceleration measures}

We have implemented a simulation to check this method of localization. We made this on the following system: $n$ masses, connected
through springs and dampers. We denote the masses $\{m_i\}_{i=1}^n$, the springs stifness $\{k_i\}_{i=1}^n$ and the dampers $\{c_i\}_{i=1}^n$ (see Figure (\ref{springs})).

@@

For the following simulation results, we consider a system with 10 masses. We put a sensor on each mass. For the damaged state, we lower the stifness of the spring between sensors 6 and 7. The values of the FRF on the figures are the norm of the FRF values.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/detect_curve_without_noise.png}
  \caption{Damage detection rate against damage (without noise)}
  \label{detect}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/detect_curve_with_005noise.png}
  \caption{}
  \label{detect_5noise}
\end{figure}

@@

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/quantification_without_noise_100.png}
  \caption{}
  \label{quant}
\end{figure}

>>>>

The purpose of this work is the analysis of a particular data-driven damage localization method,
 namely the FRF interpolation method \cite{dilena2015damage}. Our contribution is the development of a simulation to obtain acceleration measurement (which can run on several structures), a proposition of a variant of the FRF interpolation method, and the statistical analysis of the two methods.
This paper is organized as follows. In section II we  present the existing methods (model-based and data driven methods), 
then in section III we present different models of vibrating structures required for what happens next. In section IV,  we talk about the localization using FRF interpolation method, in section V we detail our simulation used to obtain data and, finally, the experiences and its results in section VI.

@@

\section{Vibration modeling}

In this section, we present the physical models that we use to obtain data for our experiences, and the Frequency Response Function (FRF) which is the base of the FRF interpolation method.

\subsection{Finite element model}

The dynamic behaviour of a mechanical system consisting of $n$ degrees of freedom (see Figure (\ref{finite_el})) is described by following matrix differential equation:
\begin{equation}
M \ddot{q}(t) + C \dot{q}(t) + K q(t) = f(t)
\label{fondamental}
\end{equation}
where $q(t)$ is the displacement vector at a continous time $t$, $f(t)$ is the excitement vector at $t$ and $M, C, K \in \mathbb{R}^{n \times n}$, are the mass, damping and stiffness matrices.

\subsection{Continous-time state-space model} %(model parameters and reduction):

\subsection{Discrete-time state-space model} %(impulses response)

The FRF correspond to the transfer function of the system. In frequency domain, if $F$ is the force that is apply on the system and $X$ is the response of this force, the FRF is given by:
\begin{equation}
H = \frac{X}{F}
\end{equation}

We can express the theoretical FRF with the matrices $M, C, K$. Indeed, in the frequency domain, we have
\begin{equation}
s^2 M x + s C x + K x = f
\end{equation}
where s is $i\omega$, with $\omega$ the pulsation.
Thus, the FRF $H$ is:
\begin{equation}
H = \frac{x}{f} = (s^2 M + s C + K)^{-1}
\end{equation}

In practice, we want to calculate the FRF of the system, for each sensor localization. Here, we calculate an estimation of the FRF, called $H_1$ estimator \cite{cauberghe2004applied}. We suppose we have a vector of the forces $\{ f_k\}_{k=1}^N$ and a vector of the accelerations $\{ y_k\}_{k=1}^N$ for each sensor (the two vectors are size $N$). We need to be in frequency domain, so we calculate the discrete Fourier transform:
\begin{equation}
\begin{array}{ll}
F_n = \sum\limits_{k=1}^N f_k e^{-2i\pi kn/N} \ \ \ \ \ n \in \{1 ... N\} \\
Y_n = \sum\limits_{k=1}^N y_k e^{-2i\pi kn/N} \ \ \ \ \ n \in \{1 ... N\}
\end{array}
\end{equation}

@@

Figure (\ref{theo}) shows this loss of spatial regularity on the theoretical FRF, for a damage at the position 7.
We can see there is a loss of regularity at the position 7 for the damaged state. We can also see there is an iregularity at the position 2, but here the two states have the same iregularity, so it's not a loss of regularity.

@@

In this section, we suppose that we have $n$ sensors, which are located along an axis $z$, and we denote them by
$\{z_l\}, l\in \{1 ... n\}$ (that is their position on the axis $z$). We also suppose we have $N$ measurements for each sensor.

For each sensor, we calculate the Frequency Response Function (FRF) from the acceleration's measures, at certain frequencies 
(which are $\frac{k}{N}, k \in \{0 ... N-1\}$ with $N$ the size of the input and output vectors).

@@

We can see this error on the Figure(\ref{interpolation_error}).

@@

@@

We have to calculate for each sensor $z_l$ and for each considered frequency $f_i$ the cubic polynomial spline interpolation of the transfer function $H_R(z_k,f_i)$.
The cubic spline interpolation is a piecewise continuous curve, passing through each of the values.
There is a separate cubic polynomial interpolation for each interval, each with its own coefficients.
This is for this local aspect that we use this interpolation, since we want to detect local iregularities.
Cubic polynomial spline interpolations produce an interpolated function that is continuous through to the second derivative.

@@

With this interpolation, we can calculate the error as it is described previously.

\subsection{Threshold}

@@

Thus, we can have different values of the error interpolation for a same state. Heowever, the distributions of the erros for a same state is close to a gaussian distribution. The problem is that the distributions for a undamaged and a damaged states can be near.

So that means once have to determine a threshold $E_T$, such as $ \Delta E(z_l)$ must be taller than $E_T$ to have a high probability of identification of a damage at location $z_l$. Indeed, as we can see in the Figure (\ref{proba}), the curves which represent the fact to have a damage or not have a share area: if the interpolation error $ \Delta E(z_l)$ is in this area, one can commit two errors:
\begin{itemize}
\item if it is in the area $P_f(z_l)$ under the graph of $p_{E,0}(z_l)$, once can detect a damage since there isn't one (false alarm).
\item if it is in the area $P_m(z_l)$ under the graph of $p_{E,d}(z_l)$, once can not detect a damage since there is one (missing alarm).
\end{itemize}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/explication_FP.png}
  \caption{Probabilities of false and missing alarm, from \cite{dilena2015damage}}
  \label{proba}
\end{figure}

\begin{figure}[h!]
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
\ & Damaged & Undamaged \\ \hline
Detected & TP & FP \\ \hline
No detected & FN & TN \\ \hline
\end{tabular}
\end{center}
\caption{True/False positives and negatives}
\label{TFPN}
\end{figure}

Now, to denote corect detection (or no detection), we will use the terms true positives and negatives, to denote missing alarms we will use false negatives and for false alarms, we will use false positives.

In our experiences, we measure the number of true/false positives and negatives. The figure(\ref{TFPN}) present the notations used for the number of true/false positives and negatives.

To evaluate the two methods, we use the specificity predictive factors, which are 
$\frac{TP}{TP + FP}$ and $\frac{TN}{TN + FN}$. The first correspond to the probability that there is a damage when we detect one, and the second correspond to the probability there isn't any damage when we detect nothing. A good method has the two factors near to 1.

@@

We propose here a variant of the FRF interpolation method. While we have worked on the implementation of the original method, we have observed the theoretical FRF to see if the FRF interpolation method worked on that. We noticed that the following method is better than the interpolation method (compute the interpolation error, then compute a difference) to detect damaged on the eigenvectors:
\begin{itemize}
\item first, compute the difference between the eigenvectors
\item second, compute the interpolation error on this difference
\end{itemize}

Figure(\ref{theodiff}) shows the difference betweem the two FRF, and we can see the loss of spatial regularity (damage at position 7).
In comparaison with the Figure(\ref{theo}), we can see here there is only one iregularity: the fact to compute the difference before the interpolation erase iregularity when they are in the two states. With the initial method, they are also erase when we compute the difference between the interpolation errors.

@@

We have implemented a simulation to obtain the data used to check the methods of localization. To explain it, we detil here the following system we used for our experiences: $n$ masses, connected
through springs and dampers. We denote the masses $\{m_i\}_{i=1}^n$, the springs stifness $\{k_i\}_{i=1}^n$ and the dampers $\{c_i\}_{i=1}^n$ (see Figure (\ref{springs})).

@@

For the following simulation results, we consider a system with 20 masses. We put a sensor on each mass. For the damaged state, we lower the stifness of the spring between sensors 6 and 7. The values of the FRF on the figures are the norm of the complex FRF values.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/detect004.png}
  \caption{Damage detection rate against damage (without noise)}
  \label{detect}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/detect_noise004.png}
  \caption{}
  \label{detect_5noise}
\end{figure}

@@

Figure (\ref{pred_max}) shows the predicitve factor with this method.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/pred_max.png}
  \caption{Predictive factor}
  \label{pred_max}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/quantification.png}
  \caption{}
  \label{quant}
\end{figure}
